### 2020-06-27 paper1/paper1.tex
Add paper1 template

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Automated Network Packet Generation for Evaluation of Neural Networks in Intrusion Prevention Systems}

\author{\IEEEauthorblockN{Bernhard Gally}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria\\
cs19m023@technikum-wien.at}
\and
\IEEEauthorblockN{Sebastian Lipp}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m032@technikum-wien.at}
\and
\IEEEauthorblockN{Damir Marijanovic}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m031@technikum-wien.at}
}

\maketitle

\begin{abstract}
@Damir
\end{abstract}

\begin{IEEEkeywords}
artificial intelligence, neural networks, computer networks, network security, packet generation
\end{IEEEkeywords}

\section{Introduction}
@Damir

\section{Infrastructure}
@Sebastian

\section{Simulation}
@Bernhard

\section{Conclusion}
@Sebastian

\begin{thebibliography}{00}
\bibitem{b1} zzz
\end{thebibliography}

\end{document}